### 2021-01-11 Manuscript_pirAnalysis_Github/AncestralPir/AnalysisCode/Manuscript_Github_AncestralPir.Rmd
Add files via upload

---
title: "Manuscript_Github_ancestral_pir_210801"
author: "Timothy Little"
date: "08/01/2021"
output: 
  html_document:
      toc: true
      number_sections: true
      toc_float: true
---

Got the 19 individual orthologs of PBANKA_0100500 from PlasmoDB, did orthology transformation twice to get everything. Then added the PccCB ortholog = 20 genes.
With the 20 genes did a MUSCLE (v3.8.31) alignment, PhyML (v3.1)* building with LG model 100 bootstrap replicates, on SEAVIEW v4.7.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = 
                       "/Users/littlet/Dropbox (The Francis Crick)/Data/Manuscript_pirAnalysis_Github/AncestralPir/AnalysisCode")
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r load_packages}
library('treeio') # Download trees from various formats
library('ggtree') # Draw phylogenetic trees with ggplot
library('ggstance')
library('ggnewscale')
library('plotly')
library('ComplexHeatmap')
library('circlize') #for ColorRamp2 function
library('ggplot2')
library('data.table')
library('gridExtra')
library('tibble')
library('stringr')
library('cowplot')
library('kableExtra')
library('viridis') # color palettes
library('RColorBrewer') # color palettes
library('corrplot')
library('reshape2')
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('readxl') # for reading in Excel spreadsheets
library('readr') # for reading in everything else
# library('tidyverse')
library('dplyr')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r function_uniquify}
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
    name <- paste(vector_names[i],
                  df_num$rep[match(vector_names[i], df_num$names)],
                  sep = '_')
    #Increase rep counter
    df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
    new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r tpm_function}
tpm = function(count_data, lengths) {

  #Print an error if the id columns do not match
  # first if statement to test that both have a GeneID column
  if(any(grepl(colnames(lengths), 
               pattern = 'id', 
               ignore.case = T)) == T & any(grepl(colnames(count_data), 
                                                  pattern = 'Geneid', 
                                                  #This needs to be Geneid when some of the samples have 'id' in them
                                                  ignore.case = T)) == T){
    if(all(count_data[grep(colnames(count_data), 
                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
                                                                             pattern = 'id', ignore.case = T)])) {
      print('Count data and length gene ids both match')
    } else {
      print('Warning:Count data and length gene ids DO NOT MATCH')
    }
  } else {
    print('Count data and/or length table supplied do not have an id column, 
          make sure they are ordered correctly')
  }

  #Remove the id column from supplied count_data
  if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = TRUE))){
    gene_names <- count_data[,grep(colnames(count_data), pattern = 'Geneid', ignore.case = TRUE)]
    counts_tpm <- count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = TRUE)]
  }else{
    counts_tpm <- count_data}

  #Remove the id column from supplied lengths
  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = TRUE))){
    lengths.table = lengths[,grep(colnames(lengths), pattern = 'length', ignore.case = T)]
  }else{
    lengths.table <- lengths}
  #Making lengths.table a matrix sometimes throws up problems
  lengths.table = as.matrix(lengths.table)
  #This may make lengths.table a vector which can bring up some dimension problems in the following for loop

  #Now for the actual calculation
  rpm <- counts_tpm
  for (i in 1:dim(counts_tpm)[2]) {
    for (j in 1:dim(counts_tpm)[1]) {
      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j]/1000))
    }
    scale.factor <- sum(rpm[,i])/(1E6)
    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
  }
  colnames(counts_tpm) <- 'tpm'
  if(any(grepl(colnames(count_data), pattern = 'Geneid', ignore.case = T)) == T){
    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), 
                                            pattern = 'Geneid', 
                                            ignore.case = T), 
                                      drop = FALSE],
                           counts_tpm)
  } 
  return(counts_tpm)
}
```

```{r function_get_ecdf_percentile}
ecdf_percentile <- function(df_tpm){
  Fn <- ecdf(
    unlist( #To convert into an atomic vector that ecdf will like.
      select_at(
        df_tpm,
        vars(
          matches('med_fpkm|med_tpm|^tpm$') #Default matches behaviour is ignore.case = TRUE
        ) #Could be med_fpkm or med_tpm.
      )
    )
  )
  #Using a lapply to make a list in case there are multiple ancestral genes.
  list_percentile <- lapply(
    df_tpm$Geneid[
      df_tpm$ancestral_ortholog == 1
      ],
    function(ancestor_gene){
      #For debugging
      # ancestor_gene <- "PCHCB_0101200"
      list(
        Fn(
          filter(df_tpm, Geneid %in% ancestor_gene) %>% 
            select_at(
              vars(
                matches('med_fpkm|med_tpm|^tpm$')
              ) #Could be med_fpkm or med_tpm
            ) 
        )
      )
    }
  )
  #Create data.frame of the output
  gene_percentile_df <- data.frame(
    Geneid = df_tpm$Geneid[
      df_tpm$ancestral_ortholog ==1
      ],
    percentile = unlist(
      lapply(
        list_percentile, `[`, 1
      )
    )
  )
  plot(Fn)
  return(gene_percentile_df)
}
# ecdf_percentile(cyno_genes_fpkm)
```

```{r get_phylip_tree}
phylog <- read.tree('PlasmoDBv48_GenesByOrthologs_ancestralpir_withPccCBandPberghei-PhyML_100bootstrap_tree.phylip')
```

```{r removing_unncessary_tips}
phylog <- drop.tip(phylog, 
                   tip = c("PKNOH_S130216000",
                           "YYE_04766",
                           "YYG_04764",
                           # "PY17X_0102100",
                           "PY06119",
                           'PYYM_0101500',
                           'PVL_110036700')
)
```

```{r gene_and_species_names}
gene_names <- phylog$tip.label %>% 
  str_split_fixed(pattern = '_P.', n =2) %>% 
  .[,1]
write(gene_names,
            file = 'gene_names_ancestral_includedintree.txt',
            sep = '\n')
phylog$old.tip.label <- phylog$tip.label

#Download the gene info with species names and add in the PccCB name.
ortho_gene_info_tbl <- rbind(
  read_csv('PlasmoDBv48_ancestralpir_orthologs_GenesByOrthologs_Summary.csv') %>% 
    select(c('Gene ID', Organism)) %>% 
    dplyr::rename(Geneid = 'Gene ID'),
  data.frame(Geneid = 'PCHCB_0101200',
             'Organism' = 'Plasmodium chabaudi chabaudi CB')
)

phylog$tip.label <- ortho_gene_info_tbl$Organism[match(phylog$old.tip.label, 
                                                       ortho_gene_info_tbl$Geneid)] %>% 
  str_replace(pattern = 'Plasmodium ', 
              replacement = 'P. ' ) %>% 
  str_replace(pattern = '(?<!cynomolgi) strain.+| curtisi|(?<=chabaudi) chabaudi|(?<=yoelii) yoelii', 
              replacement = '') %>% 
  str_extract(pattern = 'P. [[:lower:]]+(-like)?( [[:lower:]]+)?( M| B| CB| Sal-1| P01|)') %>% 
  gsub(pattern = 'chabaudi$', 
       replacement = 'chabaudi AS') %>% 
  gsub(pattern = 'strain ',
       replacement = '') %>% 
  uniquify(first_one = TRUE) 
#There shouldn't be any extra ones now
ovale_gene_names <- str_subset(gene_names, 
                               pattern = 'Poc')
```

'A glance of the blood stage transcriptome of a Southeast Asian Plasmodium ovale isolate'

ovale data is in TPM

```{r ovale_transcriptome_data}
ovale_path <- '../other_species_transcriptome_data/Brashear_2019_ovale_transcriptome_journal.pntd.0007850.s008.xlsx'
ovale <- read_xlsx(ovale_path, sheet = 1)
ovale <- ovale %>% 
  mutate(ancestral_ortholog = ifelse(Gene %in% gene_names, 1, 0))
#Sanity test
(ovale$ancestral_ortholog %>% sum) == 9

ovale_anc_v_nonanc <- ovale %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(med_tpm = median(TPM),
                   mean_tpm = mean(TPM),
                   max_tpm = max(TPM),
                   min_tpm = min(TPM))
#Median TPM of non-ancestral pirs
ovale_nonanc_medtpm <- ovale_anc_v_nonanc %>% filter(ancestral_ortholog == 0) %>% .$med_tpm
#Ratio of the TPM of the ancestrals to the median non-ancestrals
ovale_ancestral <- ovale %>% 
  filter(ancestral_ortholog ==1) %>% 
  mutate(Geneid = Gene,
         ratio = TPM/ovale_nonanc_medtpm,
         median_tpm = TPM) %>% 
  select(Geneid, ratio, median_tpm)
```

```{r ovale_percentile}
#Calculate the percentile of the ancestral v other pirs

ovale %>% 
  filter(str_detect(Description, pattern = 'PIR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

ovale[ovale$ancestral_ortholog == 1,]

ovale_tpm_collated_pir <- ovale  %>% 
  filter(str_detect(Description, pattern = 'PIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo') | ancestral_ortholog == 1) %>% 
  select(c(Gene, Description, ancestral_ortholog, TPM)) %>% 
  dplyr::arrange(desc(TPM))

#ecdf produces a function which can be used to find the percentile that a value is present in in a distribution
Fn_ovale <- ecdf(ovale_tpm_collated_pir$TPM)
plot(Fn_ovale)
list_ovale_percantile <- lapply(ovale_tpm_collated_pir$Gene[ovale_tpm_collated_pir$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_ovale(filter(ovale_tpm_collated_pir, Gene %in% ancestr)$TPM))
                                })
ovale_percentile <- data.frame(Geneid = ovale_tpm_collated_pir$Gene[ovale_tpm_collated_pir$ancestral_ortholog == 1],
                               percentile = unlist(lapply(list_ovale_percantile, `[`, 1)))

#Percentile for all genes
# Fn_ovale_all <- ecdf(ovale$TPM)
# plot(Fn_ovale_all)
# list_ovale_percantile_all <- lapply(ovale$Gene[ovale$ancestral_ortholog == 1], 
#                                 function(ancestr){
#                                   list(Fn_ovale_all(filter(ovale, Gene %in% ancestr)$TPM))
#                                 })
# (ovale_percentile_all <- data.frame(Geneid = ovale$Gene[ovale$ancestral_ortholog == 1],
#                                percentile = unlist(lapply(list_ovale_percantile_all, `[`, 1))))

#Wow must genes in the genome are expressed super low (52% expressed under 25 TPM)
# nrow(filter(ovale, TPM < 25)) /72.80

# ecdf_percentile(mutate(ovale, med_tpm = TPM, Geneid = Gene))
ovale_percentile_all <- ecdf_percentile(mutate(ovale, Geneid = Gene))

```

Data from eLife _A comparative transcriptomic analysis of replicating and dormant liver stages of the relapsing malaria parasite Plasmodium cynomolgi_ 2017.

Note: FPKM

```{r cynomolgi_transcriptome_data}
cyno_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'cynomolgi',
                        full.names = TRUE)
cyno <- read_xlsx(cyno_path, sheet = 1, skip = 3)
colnames(cyno) <- make.names(colnames(cyno))
apply(cyno %>% select_at(vars(matches('Sz|Hz'))), 2, sum)
cyno <- cyno %>% 
  mutate(ancestral_ortholog = ifelse(Gene.Id %in% gene_names,1,0))
(cyno$ancestral_ortholog %>% sum)
cyno_anc_v_nonanc <- cyno %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('Sz\\.|Hz\\.')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  #gsub to remove the numbers for easy summarising across the replicates
  mutate(stage = gsub(variable, pattern = '.[[:digit:]]', replacement = '')) %>% 
  group_by(ancestral_ortholog, stage) %>% 
  dplyr::summarise(median = median(value),
            standev = sd(value))
cyno_schiz_ratio <- filter(cyno_anc_v_nonanc, 
                           ancestral_ortholog == 1) %>% 
  filter(stage == 'Sz') %>% 
  .$median / 
  filter(cyno_anc_v_nonanc, ancestral_ortholog == 0) %>% 
  filter(stage == 'Sz') %>% 
  .$median
cyno_hyp_ratio <- filter(cyno_anc_v_nonanc, ancestral_ortholog == 1) %>% 
  filter(stage == 'Hz') %>% 
  .$median / 
  filter(cyno_anc_v_nonanc, ancestral_ortholog == 0) %>% 
  filter(stage == 'Hz') %>% 
  .$median
cyno_data <- data.frame(Geneid = cyno$Gene.Id[which(cyno$ancestral_ortholog == 1)],
                        ratio = cyno_schiz_ratio,
                        median_tpm = filter(cyno_anc_v_nonanc, ancestral_ortholog == 1) %>% 
                          filter(stage == 'Sz') %>% 
                          .$median)
```

```{r cyno_percentile}
#Calculate the percentile of the ancestral v other pirs
cyno_gene_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'PlasmoDBv46_PcynoM',
                        full.names = TRUE)
cyno_genes <- read_csv(cyno_gene_path) %>% 
  mutate(Geneid = `Gene ID`,
         Description = `Product Description`) %>% 
  select(c(Geneid, Description))

cyno_genes %>% 
  filter(str_detect(Description, pattern = 'PIR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

cyno_genes %>% 
  filter(str_detect(Description, pattern = 'pseuodgene'))

cyno_genes_fpkm <- cyno %>% 
  select_at(vars(matches('^Gene|Average'))) %>% 
  melt(id.vars = 'Gene.Id') %>% 
  group_by(Gene.Id) %>% 
  dplyr::summarise(med_tpm = median(value),
                   sd_tpm = sd(value)) %>% 
  mutate(Description = cyno_genes$Description[match(Gene.Id, cyno_genes$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Gene.Id %in% gene_names,1,0)) %>% 
  mutate(Geneid = Gene.Id)

cyno_pir_fpkm <- cyno_genes_fpkm  %>% 
  filter(str_detect(Description, pattern = 'PIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>% 
  select(c(Geneid, Description, ancestral_ortholog, med_tpm)) %>% 
  dplyr::arrange(desc(med_tpm))
cyno_pir_fpkm[which(cyno_pir_fpkm$Geneid %in% gene_names),]
cyno_pir_fpkm[which(cyno_pir_fpkm$ancestral_ortholog == 1),]

#The percentile among the pir genes
Fn_cyno <- ecdf(cyno_pir_fpkm$med_tpm)
plot(Fn_cyno)
list_cyno_percantile <- lapply(cyno_pir_fpkm$Geneid[cyno_pir_fpkm$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_cyno(filter(cyno_pir_fpkm, Geneid %in% ancestr)$med_tpm))
                                })
cyno_percentile <- data.frame(Geneid = cyno_pir_fpkm$Geneid[cyno_pir_fpkm$ancestral_ortholog == 1],
                              percentile = unlist(lapply(list_cyno_percantile, `[`, 1)))

#The percentile among all genes
Fn_cyno_all <- ecdf(cyno_genes_fpkm$med_tpm)
plot(Fn_cyno_all)
list_cyno_percentile_all <- lapply(cyno_genes_fpkm$Geneid[cyno_genes_fpkm$ancestral_ortholog == 1], 
                                   function(ancestr){
                                     list(Fn_cyno_all(filter(cyno_genes_fpkm, Geneid %in% ancestr)$med_tpm))
                                   })
cyno_percentile_all <- data.frame(Geneid = cyno_genes_fpkm$Geneid[cyno_genes_fpkm$ancestral_ortholog == 1],
                              percentile = unlist(lapply(list_cyno_percentile_all, `[`, 1)))

ecdf_percentile(cyno_genes_fpkm)

```

Vivax data from Siegal et al 2020 SciRep 'Analysis of Plasmodium vivax schizont transcriptomes from field isolates reveals heterogeneity of expression of genes involved in host-parasite interactions'.

* FPKM data. 
* Clinical isolates.

```{r}
vivax_genes_fpkm <- read_xlsx('../other_species_transcriptome_data/Siegal2020_vivax-schizont-rnaseq_41598_2020_73562_MOESM2_ESM.xlsx',
          sheet = 3,
          skip = 1,
          col_names = TRUE) %>% 
  mutate(Geneid = gene_id,
         rpkm = `mean RPKM`,
         ancestral_ortholog = ifelse(gene_id %in% gene_names,1,0)) %>% 
  select(Geneid, 
         rpkm,
         ancestral_ortholog)

# sum(vivax_genes_fpkm$ancestral_ortholog) == 1

group_by(vivax_genes_fpkm, 
         ancestral_ortholog) %>% 
  dplyr::summarise(median = median(rpkm),
                   standev = sd(rpkm))

#The percentile among all genes
Fn_vivax_all <- ecdf(vivax_genes_fpkm$rpkm)
plot(Fn_vivax_all)
list_vivax_percentile_all <- lapply(vivax_genes_fpkm$Geneid[vivax_genes_fpkm$ancestral_ortholog == 1], 
                                   function(ancestr){
                                     list(Fn_vivax_all(filter(vivax_genes_fpkm, Geneid %in% ancestr)$rpkm))
                                   })
vivax_percentile_all <- data.frame(Geneid = vivax_genes_fpkm$Geneid[vivax_genes_fpkm$ancestral_ortholog == 1],
                              percentile = unlist(lapply(list_vivax_percentile_all, `[`, 1)))

```

P. chabaudi chabaudi CB data from Lim et al, 2019.

```{r cb_transcriptome_data}
cb_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'PccCB_rnaseq',
                        full.names = TRUE)
cb <- read_delim(cb_path, delim = '\t')
apply(cb[,-1], 2, sum) #FPKM, do not add up to 1e6, although they are close
cb <- cb %>% 
  mutate(Geneid = id) %>% 
  select(-id) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
(cb$ancestral_ortholog %>% sum)
cb_anc_v_nonanc <- cb %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('CB')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
            standev = sd(value))
cb_data <- data.frame(Geneid = cb$Geneid[which(cb$ancestral_ortholog == 1)],
                      ratio = cb_anc_v_nonanc$median[which(cb_anc_v_nonanc$ancestral_ortholog == 1)]/
                        cb_anc_v_nonanc$median[which(cb_anc_v_nonanc$ancestral_ortholog == 0)],
                      median_tpm = cb_anc_v_nonanc$median[which(cb_anc_v_nonanc$ancestral_ortholog == 1)])
```

Used ftp://ftp.sanger.ac.uk/pub/project/pathogens/Plasmodium/chabaudi/CB_v2/ for the gene names

```{r cb_percentile}
#Calculate the percentile of the ancestral v other pirs
cb_gene_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'PccCB_gene',
                        full.names = TRUE)

read_delim(cb_gene_path, delim = '\t')[!(apply((apply(read_delim(cb_gene_path, delim = '\t'),
             1,
             str_extract,
             pattern = 'pseudo') %>%
         is.na),
      2, all)),] #Are there any pseudo columns? Yes : 'attributes'

cb_genes <- read_delim(cb_gene_path, delim = '\t', 
                       col_names = c('Geneid', 'Description'))

cb_genes %>% 
  filter(str_detect(Description, pattern = 'PIR|CIR|IR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

# cb_genes %>% 
#   group_by(Description) %>% 
#   dplyr::summarise(num = n()) %>% 
#   dplyr::arrange(desc(num))

cb_genes_fpkm <- cb  %>% 
  select_at(vars(matches('TCB|Geneid'))) %>% 
  melt(id.vars = 'Geneid') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_fpkm = median(value),
                   sd_fpkm = sd(value)) %>% 
  mutate(Description = cb_genes$Description[match(Geneid, cb_genes$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

cb_pir_fpkm <- cb_genes_fpkm  %>% 
  filter(str_detect(Description, pattern = 'CIR|PIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>% 
  select(c(Geneid, Description, ancestral_ortholog, med_fpkm)) %>% 
  dplyr::arrange(desc(med_fpkm))
cb_pir_fpkm[which(cb_pir_fpkm$Geneid %in% gene_names),]

Fn_cb <- ecdf(cb_pir_fpkm$med_fpkm)
plot(Fn_cb)
list_cb_percantile <- lapply(cb_pir_fpkm$Geneid[cb_pir_fpkm$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_cb(filter(cb_pir_fpkm, Geneid %in% ancestr)$med_fpkm))
                                })
cb_percentile <- data.frame(Geneid = cb_pir_fpkm$Geneid[cb_pir_fpkm$ancestral_ortholog == 1],
                            percentile = unlist(lapply(list_cb_percantile, `[`, 1)))

ecdf_percentile(cb_pir_fpkm)
cb_percentile_all <- ecdf_percentile(cb_genes_fpkm)

```

vinckei data from Ramaprasad 2019 'A fast and cost-effective microsampling protocol incorporating reduced animal usage for time-series transcriptomics in rodent malaria parasites'

PlasmoDB v48 is not updated enough to have the vinckei assembly that the paper used. By BLASTing the PlasmoDB sequence against the PvvCY assembly on https://www.ncbi.nlm.nih.gov/assembly/GCA_900681995.1 and using the gff file from ftp://ftp.ncbi.nlm.nih.gov/genomes/all/GCA/900/681/995/GCA_900681995.1_PVVCY_v1 my guess is that it is this gene:

LR215057.1	EMBL	gene	71066	72501	.	-	.	ID=gene-PVVCY_0100190;Name=PVVCY_0100190;gbkey=Gene;gene_biotype=protein_coding;locus_tag=PVVCY_0100190

corresponding CDS data:
\>lcl|LR215057.1_cds_VEV54408.1_19 \[locus_tag=PVVCY_0100190\] \[protein=PIR protein CIR protein\] \[protein_id=VEV54408.1\] [location=complement(join(71066..71140,71320..72291,72493..72501))] \[gbkey=CDS\]

Also have downloaded the features table from the same ftp site as above.

Is this TPM or FPKM? FPKM

```{r vinck_data}
vinck_path <- list.files('../other_species_transcriptome_data/', 
                      #Use glob2rx to get the regex we want
                      pattern = glob2rx('*vinckei*xlsx*'),
                      full.names = TRUE)
vinck <- read_xlsx(vinck_path, skip =1)
select_at(vinck, vars(matches('[[:digit:]]h'))) %>% apply(., 2, sum)
#Definitely FPKM
vinck <- vinck %>% 
  mutate(Geneid = gene_id) %>% 
  select(-gene_id) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
(vinck$ancestral_ortholog %>% sum)
vinck_anc_v_nonanc <- vinck %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('[[:digit:]]h')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
            standev = sd(value))
vinck_data <- data.frame(Geneid = vinck$Geneid[which(vinck$ancestral_ortholog == 1)],
                        ratio = vinck_anc_v_nonanc$median[which(vinck_anc_v_nonanc$ancestral_ortholog == 1)]/
                        vinck_anc_v_nonanc$median[which(vinck_anc_v_nonanc$ancestral_ortholog == 0)],
                        median_tpm = vinck_anc_v_nonanc$median[which(vinck_anc_v_nonanc$ancestral_ortholog == 1)])
```

```{r vinckei_percentile}
#Calculate the percentile of the ancestral v other pirs
vinck_gene_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'vinckei_feature',
                        full.names = TRUE)

read_delim(vinck_gene_path, delim = '\t')[!(apply((apply(read_delim(vinck_gene_path, delim = '\t'),
             1,
             str_extract,
             pattern = 'pseudo') %>%
         is.na),
      2, all)),] #Are there any pseudo columns? Yes : 'attributes'

vinck_genes <- read_delim(vinck_gene_path, delim = '\t') %>% 
  mutate(Geneid = locus_tag,
         Description = ifelse(is.na(attributes),
                              name,
                              paste(name, attributes))) %>% 
  select(c(Geneid, Description)) %>% 
  filter(!is.na(Description))

vinck_genes %>% filter(Geneid %in% 'PVVCY_0200040')

vinck_genes %>% 
  filter(str_detect(Description, pattern = 'PIR|CIR|IR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

vinck_genes_fpkm <- vinck %>% 
  select_at(vars(matches('^[[:digit:]]+h|Geneid'))) %>% 
  melt(id.vars = 'Geneid') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_fpkm = median(value),
                   sd_fpkm = sd(value)) %>% 
  mutate(Description = vinck_genes$Description[match(Geneid, vinck_genes$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

vinck_pir_fpkm <- vinck_genes_fpkm  %>% 
  filter(str_detect(Description, pattern = 'PIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>% 
  select(c(Geneid, Description, ancestral_ortholog, med_fpkm)) %>% 
  dplyr::arrange(desc(med_fpkm))
vinck_pir_fpkm[which(vinck_pir_fpkm$Geneid %in% gene_names),]

Fn_vinck <- ecdf(vinck_pir_fpkm$med_fpkm)
plot(Fn_vinck)
list_vinck_percantile <- lapply(vinck_pir_fpkm$Geneid[vinck_pir_fpkm$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_vinck(filter(vinck_pir_fpkm, Geneid %in% ancestr)$med_fpkm))
                                })
vinck_percentile <- data.frame(Geneid = vinck_pir_fpkm$Geneid[vinck_pir_fpkm$ancestral_ortholog == 1],
                               percentile = unlist(lapply(list_vinck_percantile, `[`, 1)))
ecdf_percentile(vinck_pir_fpkm)
vinck_percentile_all <- ecdf_percentile(vinck_genes_fpkm)

```

This distribution of gene expression is different compared to that of the other samples..

##Coatneyi data

'Distinct amino acid and lipid perturbations characterize acute versus chronic malaria'

from https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE103259

Want to convert the counts from this file into TPM with the transcript lengths.

_P. coatneyi_ _ancestral pir_ is not annotated as a _pir_ !

```{r coatneyi_data}
coat_path <- list.files('../other_species_transcriptome_data/', 
                      pattern = '2019_coatneyi_rawcounts',
                      full.names = TRUE)
coat <- read_xlsx(coat_path, skip =3, sheet  =2)
colnames(coat)[1:2] <- c('gene', 'Geneid')
coat <- coat %>% 
  select(-gene) %>% 
  arrange(Geneid) 
#apply(coat[,-1], 2, sum)
plot(apply(coat[,-1], 2, sum), apply(coat[,-1], 2, function(x) sum(x > 0)),
     main = 'coatneyi data features detected vs total counts',
     xlab = 'Total counts of sample',
     ylab = 'Total detected genes (>0 counts) per sample')
#Great ranges in the sum of features counted in each sample, should filter out low ones
coat <- coat[,c(TRUE, apply(coat[,-1], 2, sum) > 1e7)]

coat_transcript_lengths_path <- list.files('../../genomes_genes_annotations/', 
                      pattern = 'PlasmoDBv48_Pcoatneyi',
                      full.names = TRUE)
coat_transcript_lengths <- read_csv(coat_transcript_lengths_path) %>% 
  mutate(Geneid = `Gene ID`) %>% 
  select(c(Geneid, `Transcript Length`))

coat_tpm <- tpm(coat, 
                coat_transcript_lengths)

apply(coat_tpm[,-1], 
      2, 
      sum)

coat_tpm <- coat_tpm %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

(coat_tpm$ancestral_ortholog %>% sum)
coat_anc_v_nonanc <- coat_tpm %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('tpm')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
                   standev = sd(value))
coat_data <- data.frame(Geneid = coat_tpm$Geneid[which(coat_tpm$ancestral_ortholog == 1)],
                        ratio = coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 1)]/
                          coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 0)],
                        median_tpm = coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 1)])
```

```{r coatneyi_percentile}
#Calculate the percentile of the ancestral v other pirs
coat_gene_info <- read_csv(coat_transcript_lengths_path) %>% 
  mutate(Geneid = `Gene ID`) %>% 
  select(Geneid, `Product Description`)

coat_gene_info %>% 
  filter(str_detect(`Product Description`, pattern = 'KIR')) %>% 
  group_by(`Product Description`) %>% 
  dplyr::summarise(n())

coat_tpm_collated <- coat_tpm %>% 
  melt(id.vars = 'Geneid') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_tpm = median(value),
                   sd_tpm = sd(value)) %>% 
  mutate(Product = coat_gene_info$`Product Description`[match(Geneid, coat_gene_info$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

coat_tpm_collated_pir <- coat_tpm_collated  %>% 
  filter(str_detect(Product, pattern = 'KIR') | ancestral_ortholog == 1) %>% 
  dplyr::arrange(desc(med_tpm))

Fn_coat <- ecdf(coat_tpm_collated_pir$med_tpm)

plot(Fn_coat)

list_coat_percantile <- lapply(coat_tpm_collated_pir$Geneid[coat_tpm_collated_pir$ancestral_ortholog == 1], 
                               function(ancestr){
                                 list(Fn_coat(filter(coat_tpm_collated_pir, Geneid %in% ancestr)$med_tpm))
                               })
coat_percentile <- data.frame(Geneid = coat_tpm_collated_pir$Geneid[coat_tpm_collated_pir$ancestral_ortholog == 1],
                               percentile = unlist(lapply(list_coat_percantile, `[`, 1)))

ecdf_percentile(coat_tpm_collated_pir)
coat_percentile_all <- ecdf_percentile(coat_tpm_collated)
```

berghei data - using Brugat et al., 2017

Highest expressed pir gene of any kind is 'PBANKA_0201250' - a pseudogene.

```{r berghei_data}
berg_path <- list.files('../other_species_transcriptome_data/',
                        pattern = 'Expt02',
                        recursive = TRUE,
                        full.names = TRUE)
berg <- read_csv(berg_path) %>%  
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
(berg$ancestral_ortholog %>% sum)
berg_anc_v_nonanc <- berg %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(-matches('Geneid')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
            standev = sd(value))
berg_data <- data.frame(Geneid = berg$Geneid[which(berg$ancestral_ortholog == 1)],
                        ratio = berg_anc_v_nonanc$median[which(berg_anc_v_nonanc$ancestral_ortholog == 1)]/
                          berg_anc_v_nonanc$median[which(berg_anc_v_nonanc$ancestral_ortholog == 0)],
                        median_tpm = berg_anc_v_nonanc$median[which(berg_anc_v_nonanc$ancestral_ortholog == 1)])
```

```{r berghei_percentile}

berg_descr <- read_csv('../../genomes_genes_annotations/standard_files/PlasmoDb-47_beta_Pberghei_GenesByTaxon_Summary.csv') %>% 
  mutate(Geneid = `Gene ID`,
         Description = `Product Description`) %>% 
  select(c(Geneid, Description))
berg_pir <- berg %>% 
  melt(id.vars = 'Geneid') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_tpm = median(value),
            sd_tpm = sd(value)) %>% 
  mutate(Description = berg_descr$Description[match(Geneid, berg_descr$Geneid)]) %>% 
  filter(str_detect(Description, pattern = 'PIR')) %>% 
  filter(str_detect(Description, pattern = 'pseudo|fragment', negate = TRUE)) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0)) %>% 
  dplyr::arrange(desc(med_tpm))

Fn_bir <- ecdf(berg_pir$med_tpm)
plot(Fn_bir)
list_bir_percentile <- lapply(berg_pir$Geneid[berg_pir$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_bir(filter(berg_pir, Geneid %in% ancestr)$med_tpm))
                                })
bir_percentile <- data.frame(Geneid = berg_pir$Geneid[berg_pir$ancestral_ortholog == 1],
                             percentile = unlist(lapply(list_bir_percentile, `[`, 1)))

berg_percentile_all <- ecdf_percentile(
  berg %>% 
    melt(id.vars = 'Geneid') %>% 
    group_by(Geneid) %>% 
    dplyr::summarise(med_tpm = median(value),
              sd_tpm = sd(value)) %>% 
    mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
)
```

Py17XNL data from gametocytes (Hart ... Lindner et al 2019) 'Plasmodium male gametocyte development and transmission are critically regulated by the two putative deadenylases of the CAF1/CCR4/NOT complex'

```{r yoelii_data}
yeol_path <- list.files('..',
                        pattern = 'Hart_2019_Pyeolii',
                        recursive = TRUE,
                        full.names = TRUE)
yeol_counts_full <- read_xlsx(yeol_path, 
                         sheet = 2) %>% 
  mutate(Geneid = GeneID...1,
         Description = `[Product Description]`,
         Counts = `489 Average`) %>% 
  select(c(Geneid, Description, Counts)) %>% 
  arrange(Geneid)

yeol_counts_full %>%
  filter(str_detect(Description, pattern = 'PIR|CIR|IR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

yeol_counts <- select(yeol_counts_full, -Description)

yeol_lengths_path <- list.files('..',
                                pattern = 'PlasmoDBv46_Pyoelii17X_',
                                recursive = TRUE,
                                full.names = TRUE)
yeol_lengths <- read_csv(yeol_lengths_path) %>% 
  mutate(Geneid = `Gene ID`,
         length = `Transcript Length`) %>% 
  select(c(Geneid, length))
#One gene 'PY17X_0902700' has two different length transcripts - take median
yeol_lengths %>% group_by(Geneid) %>% dplyr::summarise(n = n()) %>% dplyr::arrange(desc(n))

#This gives 5074.5 so need to make sure that not having an integer count messes up things but it shouldn't
yeol_lengths <- yeol_lengths %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(length = median(length))
nrow(yeol_lengths)
nrow(yeol_counts)

yeol_tpm <- tpm(count_data = yeol_counts, lengths = yeol_lengths)
yeol_tpm <- yeol_tpm %>% 
  mutate(Description = yeol_counts_full$Description[match(Geneid, 
                                                          yeol_counts_full$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

yeol_tpm_pir <- yeol_tpm %>% 
  filter(str_detect(Description, pattern = 'PIR|YIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>% 
  dplyr::arrange(desc(tpm))

Fn_yeol <- ecdf(yeol_tpm_pir$tpm)

plot(Fn_yeol)

list_yeol_percantile <- lapply(yeol_tpm_pir$Geneid[yeol_tpm_pir$ancestral_ortholog == 1], 
                               function(ancestr){
                                 list(Fn_yeol(filter(yeol_tpm_pir, Geneid %in% ancestr)$tpm))
                               })
yeol_percentile <- data.frame(Geneid = yeol_tpm_pir$Geneid[yeol_tpm_pir$ancestral_ortholog == 1],
                               percentile = unlist(lapply(list_yeol_percantile, `[`, 1))) 
yeol_percentile

yeol_percentile_all <- ecdf_percentile(yeol_tpm)

```

```{r chabaudi_percentile}

chab_info <- read_csv('../../genomes_genes_annotations/standard_files/PlasmoDB-47_beta_Pchabaudichabaudi_GenesByTaxon_Summary.csv') %>% 
  mutate(Geneid = `Gene ID`,
         Description = `Product Description`) %>% 
  select(c(Geneid, Description))

chab_info %>% 
  filter(str_detect(Description, pattern = 'PIR|CIR|IR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

chab_tpm <- read_xlsx('../other_species_transcriptome_data/Manuscript_Frontiers_SuppInfo_chabaudi_24h.xlsx',
                      sheet = 'chab_TPM_avg_rMT') %>% 
  melt(variable.name = 'stage', value.name = 'tpm') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_tpm = median(tpm),
                sd_tpm = sd(tpm)) %>% 
  mutate(Description = chab_info$Description[match(Geneid, 
                                                       chab_info$Geneid)]) %>% 
      mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

chab_pir <- chab_tpm %>% 
  filter(str_detect(Description, pattern = 'PIR|YIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>%
  dplyr::arrange(desc(med_tpm))

Fn_chab <- ecdf(chab_pir$med_tpm)
plot(Fn_chab)
list_chab_percentile <- lapply(chab_pir$Geneid[chab_pir$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_chab(filter(chab_pir, Geneid %in% ancestr)$med_tpm))
                                })
chab_percentile <- data.frame(Geneid = chab_pir$Geneid[chab_pir$ancestral_ortholog == 1],
                            percentile = unlist(lapply(list_chab_percentile, `[`, 1)))

chab_percentile_all <- ecdf_percentile(chab_tpm)
```

## Phylogenetic Tree / Heatmap of percentile compared to all genes

```{r tree_heatmap_all-genes}

ancestral_df <- bind_rows(cyno_percentile_all,
                          ovale_percentile_all,
                          cb_percentile_all,
                          vinck_percentile_all,
                          coat_percentile_all,
                          yeol_percentile_all,
                          berg_percentile_all,
                          chab_percentile_all,
                          vivax_percentile_all)
ancestral_data <- lapply(phylog$old.tip.label, function(id){
  #If any of the pattern is in the string one will be TRUE
  ratio <- if(any(str_detect(string = paste(id), 
                             pattern = ancestral_df$Geneid))){
    index_id <- str_detect(string = paste(id), 
                           pattern = ancestral_df$Geneid)
    ancestral_df$percentile[index_id]*100
  }else{
    NA
  }
  return(list(ratio = ratio))
}) %>% 
  bind_rows
ancestral_data <- as.data.frame(ancestral_data)

rownames(ancestral_data) <- phylog$tip.label

breaks <- seq(0, 100, by = 10)

(tr <- ggtree(phylog, 
            branch.length = 'none',
            layout = 'circular') +
  geom_tiplab2(aes(angle=angle, x = x + 2.5), 
               cex = 3, 
               color='black',
               font = 3) +
  ggplot2::xlim(0,20) 
)
```

```{r heatmap_with_tree_figure3b}

p <- gheatmap(tr, 
              ancestral_data, 
              offset=.2, 
              width=.15, 
              colnames=FALSE) +
  scale_fill_viridis_c(option='C', 
                       name=substitute(
                         atop(
                           '',
                           atop(
                             '', 
                             atop(
                               atop(
                                 paste(
                                   "Percentile of ",
                                   italic('ancestral pir')
                                 ), 
                                 " expression compared"
                               ), 
                               "to all genes"
                             )
                           )
                         )
                       ), 
                       #Use substitute to ensure that the italics function is passed to the name argument
                       #atop ensures that the first line is above the word 'expression'
                       #Two atops used for every two lines 
                       # as a hacky way to ensure that the lines are closer together
                       na.value = 'white', 
                       breaks = breaks) +
  #Then alter theme as hacky way to make text size back to normal and increase margin so the empty line will fit
  theme(legend.title = element_text(size = 15,
                                    margin = margin(t = -2, 
                                                    r = 0, 
                                                    b = 0, 
                                                    l = 0,
                                                    unit = "mm")))

p
```