### 2021-04-21 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
Chapl/Aapl heatmap start

---
title: "Manuscript_Github_Pchabaudi_pir_TPMHeatmaps"
author: "Timothy Little"
date: "20/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

knitr::opts_knit$set(root.dir = 
                       ".")
#Note that the above doesn't give an error if the dir doesn't exist

options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages

library('readr')
library('dplyr')
library('tidyr')
library('ggplot2')
library('DESeq2')
library("pheatmap")
library("RColorBrewer")
library('data.table')
library('gridExtra')
library('tibble')
library('stringr')
library('kableExtra')
library('viridis')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots and colours
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('readxl') # for reading in Excel spreadsheets
library('plotly')
library('scales')
library('processx') # for downloading the plotly plots
library('cowplot')
library('ComplexHeatmap')
library('ggbiplot')
library('ggpattern')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

#Global save var - set to TRUE if you want to save figures.
global_save_var <- TRUE
```

```{r function_mround}
mround <- function(x,base){ 
  base*ceiling(x/base) 
} 
```

```{r function_my_max}
my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) #max robust to NAs
```

```{r function_ordering}
ordering <- function(to_order, order_vector){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = TRUE)
}

order_vector <- c("Early.Rings",
                  "Early.Mid.Rings",
                  "Late.Mid.Rings",
                  "Late.Rings",
                  "Rings.Troph.conversion",
                  "Early.Trophs",
                  "Mid.Trophs",
                  "Late.Trophs")
```

The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome. _cir_ genes are the names of _pir_ genes in _P chabaudi_.

```{r cir_id, message=FALSE, results='hide'}

# Getting the cir gene names and transcript lengths needed for calculating the transcript-per-million

chabaudi_gene_info <- read.csv('../standard_files/PlasmoDB-47_beta_Pchabaudichabaudi_GenesByTaxon_Summary.csv',
                              header = TRUE,
                              stringsAsFactors = FALSE)

chabaudi_transcript_lengths <- select(chabaudi_gene_info, c(Gene.ID, Transcript.Length)) %>% 
  arrange(Gene.ID)

chabaudi_gene_info <- dplyr::rename(chabaudi_gene_info, 
                             Geneid = Gene.ID,
                             Description = Product.Description)

```

```{r cir_info}

cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
           delim = '\t')
cir_id <- cir_info$Geneid

cir_clade_info <- read_delim('../standard_files/Pchabaudi_pir_clades.txt',
            delim = '\t')

#Add chapl_aapl numbers for easier heatmap drawing later.
cir_info <- cir_info %>% 
  group_by(chapl_aapl_name) %>% 
  mutate(chapl_aapl_name_num = n_distinct(Geneid)) %>% 
  mutate(chapl_aapl_name_num = paste0(
                       chapl_aapl_name, 
                       ' (n = ',
                       chapl_aapl_name_num,
                       ')'
                     ))
```

```{r times_stages_labels}
#Use the RMT stage differential count in order to name the samples, but keep time to retain order
times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
                                 stage_diff_count = c('Late-Rings',
                                                      'Rings-Troph-conversion',
                                                      'Early-Trophs',
                                                      'Mid-Trophs',
                                                      'Late-Trophs',
                                                      'Early-Rings',
                                                      'Early-Mid-Rings',
                                                      'Late-Mid-Rings'))
# R will convert the dashes to '.' at some point so let's have this notation here also
times_stages_table$r_stage_diff_count <- str_replace_all(times_stages_table$stage_diff_count, 
                                                         pattern = '-', 
                                                         replacement = '\\.')
```

```{r read_TPM}
#Need to use the calculateTPM script first.

twentyfour_rMT <- read_csv('Pchabaudi_TPM_allBioReps.csv')

twentyfour_avgtpm <- list(
  'RMT' = list(
    tpm = read_csv('Pchabaudi_TPM_average.csv'),
    experiment = 'RMT'
  )
)

```

```{r colours_clades_phase}
SorL_col <- setNames(
  c(    '#F0E442',
        '#56B4E9', #L
        '#E69F00', #S
        '#999999'),
  sort(
    unique(cir_info$SorL)
  )
)

SorL_col_num <- SorL_col
names(SorL_col_num) <- sapply(names(SorL_col), 
                              function(clade_name) {
                                paste0(
                                  clade_name, 
                                  ' (n = ',
                                  sum(cir_info$SorL == clade_name),
                                  ')'
                                )
                              })

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9', #L1
    '#009E73',
    '#0072B2',
    '#E69F00', #S1
    '#D55E00',
    '#CC79A7',
    '#999999'),
  sort(
    unique(cir_info$subfam)
  )
)

clade_col_num <- clade_col
names(clade_col_num) <- sapply(names(clade_col), 
                               function(clade_name) {
                                 paste0(
                                   clade_name, 
                                   ' (n = ',
                                   sum(cir_info$subfam == clade_name),
                                   ')'
                                 )
                               })

chapl_aapl_name_col <- setNames(
  c('#F0E442',
    '#56B4E9', #L1
    '#009E73',
    '#0072B2',
    '#E69F00', #S1
    '#D55E00',
    '#CC79A7',
    '#999999'),
  sort(
    unique(cir_info$chapl_aapl_name)
  )
)

chapl_aapl_name_col_num <- chapl_aapl_name_col
names(chapl_aapl_name_col_num) <- sapply(
  names(chapl_aapl_name_col), 
  function(clade_name) {
    paste0(
      clade_name, 
      ' (n = ',
      sum(cir_info$chapl_aapl_name == clade_name, 
          na.rm = TRUE),
      ')'
    )
  })
```

_pir_ genes selected by being expressed above 1 TPM in at least one sample.

Adding in housekeeping genes for comparison:
HSP70 putative  PCHAS_0721000
MSP1 PCHAS_0831300
AMA1 PCHAS_0931000

```{r highest_expressed_stage_heatmap_fig3b}

housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
                                 'MSP1' = 'PCHAS_0831300',
                                 'AMA1' = 'PCHAS_0931000')

cir_housekeeping_id <- c(cir_id, 
                         'PCHAS_0721000',
                         'PCHAS_0831300',
                         'PCHAS_0931000')

# Get order of pir genes by highest expressed genes in each sample.

combined_data_pir <- twentyfour_avgtpm$RMT$tpm %>% 
  filter(Geneid %in% cir_housekeeping_id) %>% 
  column_to_rownames(var = 'Geneid') %>% 
  filter_all(
      any_vars(. > 1)
  )

#Ordering that a more understandable order

combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
                                                   order_vector)),
                                       drop = FALSE]

combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
                                means = apply(combined_data_pir, 1, mean),
                                subfam = factor(
                                  ifelse(rownames(combined_data_pir) %in% cir_id,
                                         cir_info$subfam[match(rownames(combined_data_pir), 
                                                               cir_info$Geneid)],
                                         'ref')
                                )) %>% 
  mutate(subfam =
           forcats::fct_shift(
             relevel(
               as.factor(subfam),
               ref = 'ref')
           )
  ) %>% 
  # filter(means > 1) %>% 
  arrange(subfam, 
          desc(means)) %>% 
  .$Geneid

#Order the heatmap data by the order defined above (Subfam, then mean expression).
combined_data_pir_ht <- combined_data_pir[
  order(
    match(rownames(combined_data_pir),
          combined_data_ordered),
    na.last = NA #Remove NAs
  ),
]

#Change the colnames to the times
colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
                                              pattern = '[[:digit:]]{2}h')

#Order by time point
combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                             drop = FALSE]

#Add a column for the subfam (ref for the reference genes)
combined_data_pir_ht$subfam <- ifelse(rownames(combined_data_pir_ht) %in% cir_id,
                                      cir_info$subfam[match(rownames(combined_data_pir_ht), 
                                                            cir_info$Geneid)],
                                      'ref')

#Get the ref gene actual names instead of the Geneid
rownames(combined_data_pir_ht) <- ifelse(
  rownames(combined_data_pir_ht) %in% housekeeping_genes[1,],
  colnames(housekeeping_genes)[ match(rownames(combined_data_pir_ht), housekeeping_genes[1,]) ],
  rownames(combined_data_pir_ht)
)

#Get the colour range
maxcol <- ceiling(
  max(
    log2(1+combined_data_pir_ht[!colnames(combined_data_pir_ht) %in% 'subfam'])
  ))

# pir_highest_tpm_order <- rownames(combined_data_pir_ht)

#Sub-family annotation
clade_info <- combined_data_pir_ht$subfam
annotation_col <- c(clade_col, 'ref' = 'black')

#Subfamily annotation

clade_ha <- rowAnnotation(
  'sub-family' = anno_simple(
    clade_info,
    col = annotation_col,
  ),
  annotation_name_gp = gpar(fontsize = 10)
)

#Subfamily name labels
num_subfam <- cir_info %>% 
  group_by(subfam) %>% 
  summarise(num = n_distinct(Gene.ID))

labels_subfam <- c(setNames(paste0(num_subfam$subfam, ' (n = ', num_subfam$num, ')'),
                            num_subfam$subfam),
                   'ref' = 'ref')

anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
                                        labels = 'ancestral',
                                        labels_gp = gpar(fontsize = 7),
                                        link_width = unit(3, "mm")))

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'subfam']),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  row_title = NULL,
  left_annotation = clade_ha,
  right_annotation = anc_ha,
  row_split = forcats::fct_shift(
    relevel(
      as.factor(combined_data_pir_ht$subfam),
      ref = 'ref')
  ),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  )
)

clade_lgd <- Legend(
  title = 'sub-family',
  at = c(labels_subfam),
  legend_gp = gpar(fill = annotation_col)
)

draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     annotation_legend_side = 'right',
     annotation_legend_list = list(clade_lgd),
     merge_legend = TRUE)
```

# Supp figure 9 - Pir expression of AAPLs and ChAPLs.

_pir_ genes selected by being expressed above 1 TPM in at least one sample.

```{r chapl_appl_heatmap_suppfig9}

# Get order of pir genes by highest expressed genes in each sample.

combined_data_pir <- twentyfour_avgtpm$RMT$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  mutate(chapl_aapl_name = cir_info$chapl_aapl_name_num[match(Geneid, cir_info$Geneid)]) %>% 
  filter(!str_detect(chapl_aapl_name, pattern = 'NA')) %>% 
  column_to_rownames(var = 'Geneid') %>% 
  filter_all(
      any_vars(. > 1)
  )

#Ordering that a more understandable order

combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
                                                   order_vector)),
                                       drop = FALSE]

combined_data_ordered <- tibble(Geneid = rownames(combined_data_pir),
                                means = apply(combined_data_pir, 1, mean),
                                chapl_aapl_name = factor(
                                  combined_data_pir$chapl_aapl_name
                                )) %>% 
  arrange(chapl_aapl_name, 
          desc(means)) %>% 
  .$Geneid

#Order the heatmap data by the order defined above (chapl_aapl_name, then mean expression).
combined_data_pir_ht <- combined_data_pir[
  order(
    match(rownames(combined_data_pir),
          combined_data_ordered),
    na.last = NA #Remove NAs
  ),
]

#Change the colnames to the times
colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
                                              pattern = '[[:digit:]]{2}h|chapl_aapl_name')

#Order by time point
combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                             drop = FALSE]

#Get the ref gene actual names instead of the Geneid
rownames(combined_data_pir_ht) <- ifelse(
  rownames(combined_data_pir_ht) %in% housekeeping_genes[1,],
  colnames(housekeeping_genes)[ match(rownames(combined_data_pir_ht), housekeeping_genes[1,]) ],
  rownames(combined_data_pir_ht)
)

#Get the colour range
maxcol <- ceiling(
  max(
    log2(1+combined_data_pir_ht[!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name'])
  ))

# pir_highest_tpm_order <- rownames(combined_data_pir_ht)

#Sub-family annotation
chapl_aapl_name_info <- combined_data_pir_ht$chapl_aapl_name
annotation_col <- c(chapl_aapl_name_col_num, 'ref' = 'black')

#chapl_aapl_name annotation

chapl_aapl_name_ha <- rowAnnotation(
  'sub-family' = anno_simple(
    chapl_aapl_name_info,
    col = annotation_col,
  ),
  annotation_name_gp = gpar(fontsize = 10)
)

#chapl_aapl_name name labels
num_chapl_aapl_name <- cir_info %>% 
  group_by(chapl_aapl_name) %>% 
  summarise(num = n_distinct(Gene.ID))

labels_chapl_aapl_name <- setNames(paste0(num_chapl_aapl_name$chapl_aapl_name, 
                                          ' (n = ', num_chapl_aapl_name$num, ')'),
                                   num_chapl_aapl_name$chapl_aapl_name)

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  row_title = NULL,
  left_annotation = chapl_aapl_name_ha,
  row_split = 
    factor(combined_data_pir_ht$chapl_aapl_name,
              levels = ordering(combined_data_pir_ht$chapl_aapl_name,
                                order_vector = 'ChAPL')),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  )
)

chapl_aapl_name_lgd <- Legend(
  title = 'sub-family',
  at = factor(labels_chapl_aapl_name,
              ordering(labels_chapl_aapl_name,
                order_vector = 'ChAPL')),
  legend_gp = gpar(fill = annotation_col)
)

draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     annotation_legend_side = 'right',
     annotation_legend_list = list(chapl_aapl_name_lgd),
     merge_legend = TRUE)
```

```{r Supp_fig_AapL-ChaAPl_expression}

cir_plot <- twentyfour_rMT %>% 
  filter(Geneid %in% cir_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,2],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
    chapl_aapl = factor(cir_info$chapl_aapl_name[match(Geneid, 
                                                       cir_info$Geneid)],
                        levels = sort(unique(cir_info$chapl_aapl_name[match(Geneid, 
                                                                            cir_info$Geneid)])))) %>% 
  filter(!(time %in% c('sch', 
                       'Unknown')),
         !is.na(chapl_aapl) ) %>% 
  group_by(Geneid) %>% 
    arrange(desc(tpm))

  cir_plot_order <- cir_plot %>% 
    filter(time == '14h') %>% 
    arrange(desc(tpm))

  cir_plot <- mutate(cir_plot, 
                     Geneid = factor(Geneid, 
                                     levels = unique(cir_plot_order$Geneid)))
  cir_plot_chapl_aapl_supp_info <- cir_plot

  max_aapl_chapl <- ceiling(max(log2(cir_plot$tpm+1)))

  breaks = seq(0, max_aapl_chapl, by = 1)
    limits = c(0, max_aapl_chapl)
    point_size = 0.25

  lapply(
    sort(unique(cir_plot$time)),
    function(time_vec){

    cir_plot <- filter(cir_plot, 
                       time == time_vec)

      cir_plot$tpm <- log2(1+cir_plot$tpm)

    cir_plot$Geneid <- factor(
      str_remove(cir_plot$Geneid, 
                 pattern = 'PCHAS_'),
      levels = str_remove(levels(cir_plot$Geneid), 
                          pattern = 'PCHAS_')
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        y = tpm
      )
    ) +
      stat_summary(fun.y = median, 
                   fun.ymin = median, 
                   fun.ymax = median,
                   geom = "bar", 
                   width = 0.75, 
                   size = 0.25, 
                   position = 'dodge',
                   fill = 'lightblue') +
      geom_point(stat = 'identity', 
                 position = position_dodge(width = 0.75),
                 size = point_size) +
      scale_y_continuous(breaks = breaks,
                         limits = limits) +
      scale_fill_manual(values = c('skyblue','red')) +
      ylab('log(TPM +1)') +
      facet_grid(.~chapl_aapl, 
                 scales = "free_x",
                 space = "free_x") +
      theme_classic() +
      theme(
        axis.text.x = element_text(angle = 90,
                                   size = 5)
      ) +
      ggtitle(paste(unique(cir_plot$time)))

    if(global_save_var){
      ggsave2(
        paste0('Pchabaudi_sibe-by-side_chapl_aapl_',
               time_vec,
               '.png'),
        plot = y,
        dpi = 300,
        width = 25,
        height = 10,
        units = 'cm'
      )
    }

    return(y)
    })
```