### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
r functions moved to new folder

---
title: "P.berghei pir analysis"
author: "Timothy Little"
date: "12/03/2021"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)
# Set below to your own working directory.
knitr::opts_knit$set(root.dir = 
                       ".")
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r load_packages, include = FALSE, message = FALSE}
library('readxl')
library('readr')
library('dplyr')
library('purrr')
library('ggplot2')
library('DESeq2')
library("RColorBrewer")
library('colorspace')
library('data.table')
library('gridExtra')
library('tibble')
library('scales')
library('stringr')
library('kableExtra')
library('viridis')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('plotly')
library('ggpattern')
library('ComplexHeatmap')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, select, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

```

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
```

```{r function_mround}
#Make nicer numbers rounded up, useful for making scales for a plot.
mround <- function(x,base){ 
  base*ceiling(x/base) 
} 
```

```{r function_my_max}
#max robust to NAs
my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) 
```

```{r function_firstup}
#Capitalize first letter
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}
```

```{r ordering_vectors}
order_vector <- c('Sporo', 'Liv-pre4h', 'Liv-pre47h', 'Liv-post48h',
                 'Liv-2h','Liv-4h','Liv-12h', 'Liv-18h',
                 'Liv-24h' ,'Liv-36h','48h','54h','60h',
                 'Mero','Asex','Schiz','Gam','-Pre3h','BldMl-24h','Ook')

r_order_vector <- str_replace_all(order_vector, 
                                  pattern = '-', 
                                  replacement = '\\.')
```

```{r pir_id, message=FALSE, results='hide'}

# Getting the pir gene names and transcript lengths needed for calculating the transcript-per-million

# First we need to get the _pir gene names. These have been downloaded using a search of PlasmoDB P. berghei ANKA genome v3 using  "\*pir\*" search and manual removal of genes which clearly were not pir.

allgene.info.file <- '../standard_files/PlasmoDb-47_beta_Pberghei_GenesByTaxon_Summary.csv'
allgene.info <- as_tibble(
  read_csv(allgene.info.file)
) %>%
  mutate(Geneid = `Gene ID`, 
         Description = `Product Description`,
         is_pseudo = `Is Pseudo`,
         transcript_length = `Transcript Length`,
         paralog_count = `Paralog count`) %>% 
  select(Geneid, 
         Description, 
         is_pseudo, 
         transcript_length, 
         paralog_count)

pir_info <- allgene.info %>% 
  filter(str_detect(Description, 
                    pattern = 'PIR(?!.+pseud)'), 
         is_pseudo == 'No') %>% 
  filter(!Geneid %in% 'PBANKA_0524600')
#Removing the pir-like gene.

pir_id <- pir_info$Geneid

#pir

pir_genes_list <- lapply(
  unique(pir_info$Description), 
  function(pir_group){
    select(
      filter(
        pir_info, 
        Description %in% pir_group
      ),
      Geneid
    )
  }
)
names(pir_genes_list) <- unique(pir_info$Description)

#How many pirs are present on each chromosome?

pir_tbl = as_tibble(data.frame(Geneid = pir_id))
pir_chrom_tbl = pir_tbl %>% 
  mutate(chrom = unlist( #Getting the chromosomes by a complicated set of lapply and sapplys...
    sapply(
      lapply(
        sapply(
          as.character(pir_tbl$Geneid), 
          strsplit,
          split = ''
        ), 
        '[', 
        8:9
      ), 
      paste, 
      collapse = ''
    )
  )
  ) 
pir_chrom_tbl %>% 
  group_by(chrom) %>% 
  dplyr::summarise(number_of_pirs_in_each_chromosome = n(),
                   percentage_of_all_pirs = round(n()/nrow(pir_tbl), 3)*100) %>% 
  dplyr::arrange(desc(percentage_of_all_pirs))

#Getting the loci information
pir_loci <- data.frame(
  Geneid = pir_id,
  chrom = str_extract(pir_id, 
                      pattern = '(?<=_)[[:digit:]]{2}'),
  number = as.numeric(str_extract(pir_id, 
                                  pattern = '(?<=_[[:digit:]]{2})[[:digit:]]+'))
) %>% 
  filter(!chrom %in% '00') %>% #Remove chrom00 (contigs)
  group_by(chrom) %>% 
  mutate(jump = ifelse(lag(number, 
                           default = 1001) < 1000 & number - lag(number, 
                                                                 default = 0) > 1000,
                       1,
                       0)) %>% 
  mutate(
    jump = ifelse(cumsum(jump) == 1 | jump == 1, 1, 0)
  ) %>% 
  mutate(
    locus = paste0(
      chrom,
      ifelse(jump == 0 & number < 1000, 
             'Left',
             'Right')
    )
  )

# Detect 'jumps' of the gene number over 1000 within each chromosome, setting the lag default as 1001 in the
#   first part of the ifelse statement so that chrom2 doesn't have an erroneous jump as it starts at >1000
# Later cumsum is do that every gene after the 'jump' is included allowing us to define telomere ends.

pir_clade <- read_delim('../standard_files/Pberghei_pir_clades.txt', 
                        delim = '\t')
pir_clade <- mutate(pir_clade, 
                    locus = pir_loci$locus[match(Geneid, pir_loci$Geneid)])
pir_id <- pir_clade$Geneid

pir_clade_list <- lapply(unique(pir_clade$clade), function(clade_name){
  pir_clade$Geneid[pir_clade$clade %in% clade_name]
})
names(pir_clade_list) <- as.character(unique(pir_clade$clade))
```

```{r transcript_redundancy, message = FALSE, results='hide'}

#Note that there is some redundancy
two.transcript <- allgene.info %>% 
  dplyr::count(Geneid) %>% 
  filter(n > 1)

#Nine genes have two transcripts - need to give them one average of the transcripts since featurecounts has only
# calculated one count for the locus
two.transcript.info <- allgene.info[allgene.info$Geneid %in% two.transcript$Geneid,]
two.transcript.mean = aggregate(transcript_length~Geneid, data=two.transcript.info, 
                                FUN=function(x) c(mean=mean(x)))

#Remove the first match
allgene.info <- allgene.info[-(match(two.transcript.mean$Geneid, 
                                     allgene.info$Geneid)),]

#For the second match replace the transcript value with the mean calculated with aggregate
allgene.info$transcript_length <- as.double(allgene.info$transcript_length)
allgene.info[match(two.transcript.mean$Geneid, 
                   allgene.info$Geneid),]$transcript_length <- two.transcript.mean$transcript_length
lengths.transcript <- subset(allgene.info, select = c(Geneid, transcript_length), drop = F)
```

## Read in counts and calculate TPM

```{r read_files}
#Reading in the counts from file
berghei_counts <- read_xlsx('Manuscript_Frontiers_berghei_counts.xlsx')

berghei_tpm <- tpm(berghei_counts,
                   lengths = lengths.transcript)

apply(berghei_tpm[-1], 2, sum)

#Get a list of df for each experiment.

expt_nams <- unique(
  na.omit(
    str_extract(colnames(berghei_tpm), 
                pattern = 'E[[:digit:]]{2}')
  )
)

expt_tpms <- lapply(
  expt_nams,
  function(expt_nam){
    expt_df <- select(
      berghei_tpm,
      matches(paste0('Geneid|', expt_nam))
    ) %>% 
      rename_with(~ str_remove(.x, 'E[[:digit:]]{2}_'),
                  matches('E[[:digit:]]{2}'))
    list(tpm = expt_df,
         experiment = expt_nam)
  }
)
names(expt_tpms) <- expt_nams
```

```{r remove_E06-6h}
#Remove the Caldelari (E06) 6h samples, as they did in their paper.

expt_tpms$E06$tpm <- select_at(expt_tpms$E06$tpm,
                                          vars(-matches('Liv-6h')))
```

Going to combine certain samples based on their distinctive profiles, e.g. <4h, 6-48h, >48h liver stages.

```{r combining_samples}
names_expt_tpms <- names(expt_tpms)

expt_tpms <- lapply(names(expt_tpms), function(expt){
  col_stage <- colnames(
    getElement(
      getElement(
        expt_tpms, 
        paste(expt)
      ),
      'tpm'
    )
  )

  liver_stage <- col_stage[
    str_detect(
      col_stage,
      pattern = 'Liv-[[:digit:]]{1,2}h'
    )
    ]

  new_liver_stage <- case_when(
    str_detect(liver_stage, pattern = 'Liv-(2|4)h') ~ 'Liv-pre4h',
    # str_detect(liver_stage, pattern = 'Liv-([5-9]|[1-3][0-9]|4[0-7])h') ~ 'Liv-pre47h',
    str_detect(liver_stage, pattern = 'Liv-(4[8-9]|[5-9][0-9])h') ~ 'Liv-post48h',
    TRUE ~ as.character(liver_stage),
  )

  if(length(new_liver_stage) > 0){
    new_liver_stage <-uniquify(new_liver_stage)
  }

  col_stage[
    str_detect(
      col_stage,
      pattern = 'Liv-[[:digit:]]{1,2}h'
    )
    ] <- new_liver_stage

  bloodmeal_stage <- col_stage[
    str_detect(
      col_stage, 
      pattern = 'BldMl-(1h|3h)_[[:digit:]]+'
    )
    ]

  bloodmeal_stage <- str_replace(
    bloodmeal_stage, 
    pattern = '(1h|3h)_[[:digit:]]+', 
    replacement = 'Pre3h'
  )
  if(length(bloodmeal_stage)>0){
    bloodmeal_stage <- uniquify(bloodmeal_stage)
  }

  col_stage[
    str_detect(
      col_stage, 
      pattern = 'BldMl-(1h|3h)_[[:digit:]]+'
    )
    ] <- bloodmeal_stage

  tpm_element <- getElement(
    getElement(
      expt_tpms, 
      paste(expt)
    ),
    'tpm'
  )

  colnames(tpm_element) <- col_stage

  list(tpm = tpm_element,
       experiment =   getElement(
         getElement(
           expt_tpms,
           paste(expt)
         ),
         'experiment'
       )
  )

})
names(expt_tpms) <- names_expt_tpms

#Re-ordering the list
expt_tpms <- (expt_tpms[order(names(expt_tpms))])
```

```{r hclusts, message = FALSE}

hclusts = function(tpm){
  if(ncol(tpm$tpm) > 3){
    dists = dist(t(tpm$tpm[,-1]))
    par(cex=0.7)
    plot(hclust(dists),
         main = tpm$experiment)
  }
}

lapply(expt_tpms, hclusts)
```

```{r remove_badly_clustering_samples}
expt_tpms$E08$tpm$`Liv-pre4h_1` <- NULL
expt_tpms$E03$tpm$`Asex-Schiz_1` <- NULL
expt_tpms$E03$tpm$`Asex-Schiz_3` <- NULL
```

## Saving the TPM

```{r save_tpm}

```

## The highest expressed _pir_ genes across the samples

```{r top_genes, echo = FALSE, message = FALSE, results = "asis"}
top_n_genes <- function(tpm_input, 
                       n = 5, 
                       threshold = 0){
  tpm = tpm_input$tpm %>% 
    #Melt the data frame for using dplyr
    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
    mutate(stage = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,1))) %>% 
    group_by(Geneid, stage) %>%  
    dplyr::summarise(median_tpm = median(tpm),
                     mean_tpm = mean(tpm), 
                     min_tpm = min(tpm), 
                     max_tpm = max(tpm), 
                     sd_tpm = sd(tpm)) %>% 
    group_by(stage) %>% 
    mutate(experiment = tpm_input$experiment)

  #Calculating the index used to make the kable table

  kable_index = rep(n, times = length(unique(tpm$stage)))
  names(kable_index) = unique(tpm$stage)

  all_genes = tpm %>% 
    top_n(wt = mean_tpm, n = n) %>% 
    dplyr::arrange(stage, desc(mean_tpm)) %>% 
    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
                                color = ifelse(mean_tpm > threshold, "red", "blue")))
  all_genes$short_descr = allgene.info$handy_descriptions[match(all_genes$Geneid, 
                                                                allgene.info$Geneid)]
  all_genes$description = allgene.info$Description[match(all_genes$Geneid, 
                                                         allgene.info$Geneid)]
  all_genes = all_genes %>% 
    kable(format = 'html', 
          caption = paste0(paste0(paste0(tpm_input$experiment, ' - Top '), paste(n)), ' expressed total genes'), 
          escape = FALSE) %>%
    kable_styling() %>% 
    pack_rows(index = kable_index)

  pir_genes = tpm %>% 
    filter(Geneid %in% pir_id) %>% 
    top_n(wt = mean_tpm, n = n) %>% 
    dplyr::arrange(stage, desc(mean_tpm)) %>% 
    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
                                color = ifelse(mean_tpm > threshold, "red", "blue"))) %>% 
    mutate(Geneid = as.character(Geneid)) %>% 
    mutate(Description = allgene.info$handy_descriptions[match(Geneid, 
                                                               allgene.info$Geneid)]) %>% 
    mutate(Geneid = ifelse(grepl(Geneid,
                                 pattern = 'PBANKA_0100500'),
                           cell_spec(Geneid, bold = TRUE),
                           Geneid)) %>%
    kable(format = 'html', 
          caption = paste0(paste0(paste0(tpm_input$experiment, ' - Top '), paste(n)), ' expressed pir genes'), 
          escape = FALSE) %>% 
    pack_rows(index = kable_index) %>% 
    kable_styling()

  # print(paste0('Completed analysis of ', tpm_input$experiment))
  list = list(tpm, all_genes, pir_genes, tpm_input$experiment)
  names(list) = c('tpm', 'all_genes', 'pir_genes', 'experiment')
  return(list)
}

### Generate and save top genes - core dataset all_top

all_top <- lapply(expt_tpms, 
                  top_n_genes)

lapply(all_top, `[`, c('all_genes', 'pir_genes'))

```

##Supp figure 1A - Pearson correlation plot of collated samples.

```{r all_samples_collated_heatmap}

big_tpm <- bind_cols(
  lapply(
    expt_tpms,
    function(x){
      tpm_df <- x$tpm
      colnames(tpm_df)[-1] <- paste0(x$experiment,
                                     '_',
                                     colnames(tpm_df)[-1])
      tpm_df
    }
  )
) %>% 
  dplyr::rename(Geneid = Geneid...1) %>% 
  select_at(vars(-matches('Geneid\\.\\.\\.\\d+$'))) %>% 
  data.frame

rownames(big_tpm) <- big_tpm$Geneid

big_tpm <- big_tpm[,-1,drop=FALSE]

## Collated across the stages for a more easily understood heatmap

big_top = bind_rows(lapply(all_top,function(x){
  data = x$tpm
  data$stage = paste(x$experiment, data$stage, sep = '_')
  data
}))

big_top_recast = dcast(big_top, Geneid ~ stage, value.var = 'mean_tpm')
big_top_recast <- column_to_rownames(big_top_recast, var = 'Geneid')
# big_top_recast = big_top_recast[,-1,drop=FALSE]

# apply(big_top_recast, 2, sum)

pdf(file = 'Manuscript_suppfig1A_corrplot_all_experiments_rnaseq.pdf',
    height = 5,
    width = 5)
corrplot(cor(big_top_recast), method = 'color', 
         order = 'hclust', hclust.method = 'ward.D2',
         cl.lim = c(-0.036,1), tl.cex = 0.50,
         col = c(viridis(100), magma(115)), addrect = 6)
dev.off()

corrplot(cor(big_top_recast), method = 'color', 
         order = 'hclust', hclust.method = 'ward.D2',
         cl.lim = c(-0.036,1), tl.cex = 0.50,
         col = c(viridis(100), magma(115)), addrect = 6)
```

## Supp figure 1B - PCA of all the samples.

```{r supp_figure_PCA_allgenes}

big_tpm_pca <-  prcomp(big_tpm, 
                       center = TRUE, 
                       scale. = TRUE)

big_tpm_pca_var <- get_pca_var(big_tpm_pca)
big_tpm_pca_coords <- data.frame(sample = rownames(big_tpm_pca_var$coord),
                                 big_tpm_pca_var$coord)
big_tpm_pca_coords$stage_experiment <- str_extract(
  big_tpm_pca_coords$sample,
  pattern = '.+(?=_[[:digit:]]+$)'
)
big_tpm_pca_coords$stage <- str_extract(big_tpm_pca_coords$sample, 
                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
big_tpm_pca_coords$experiment <- str_extract(big_tpm_pca_coords$sample, 
                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')

colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
                               length.out = length(unique(big_tpm_pca_coords$stage))),
                       unique(big_tpm_pca_coords$stage))

pca_importance <- as.data.frame(summary(big_tpm_pca)$importance)

plot <- ggplot(big_tpm_pca_coords,
       aes(
         x = Dim.1,
         y = Dim.2,
         col = stage,
         shape = experiment
       )) +
  geom_point() +
  scale_fill_manual(values = colornames) +
  scale_shape_manual(
    values = setNames(
      c(1:10),
      unique(big_tpm_pca_coords$experiment)
    )
  ) +
  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
  guides(shape=guide_legend(ncol=2),
         col = guide_legend(ncol=2)) +
  theme_classic() +
  theme(legend.key.size = unit(5, 'mm'),
        legend.text = element_text(size = 6))

plot

```

#Supp figure 1C - PCA using just the pir genes.

```{r supp_figure_PCA_pir}
big_tpm_pir <- big_tpm[rownames(big_tpm) %in% pir_id,,drop = FALSE]

big_tpm_pir_pca <-  prcomp(big_tpm_pir, 
                       center = TRUE, 
                       scale. = TRUE)

big_tpm_pir_pca_var <- get_pca_var(big_tpm_pir_pca)
big_tpm_pir_pca_coords <- data.frame(sample = rownames(big_tpm_pir_pca_var$coord),
                                 big_tpm_pir_pca_var$coord)
big_tpm_pir_pca_coords$stage_experiment <- str_extract(
  big_tpm_pir_pca_coords$sample,
  pattern = '.+(?=_[[:digit:]]+$)'
)
big_tpm_pir_pca_coords$stage <- str_extract(big_tpm_pir_pca_coords$sample, 
                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
big_tpm_pir_pca_coords$experiment <- str_extract(big_tpm_pir_pca_coords$sample, 
                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')

colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
                               length.out = length(unique(big_tpm_pir_pca_coords$stage))),
                       unique(big_tpm_pir_pca_coords$stage))

pca_importance <- as.data.frame(summary(big_tpm_pir_pca)$importance)

plot_pir <- ggplot(big_tpm_pir_pca_coords,
       aes(
         x = Dim.1,
         y = Dim.2,
         col = stage,
         shape = experiment
       )) +
  geom_point() +
  scale_fill_manual(values = colornames) +
  scale_shape_manual(
    values = setNames(
      c(1:10),
      unique(big_tpm_pca_coords$experiment)
    )
  ) +
  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
  guides(shape=guide_legend(ncol=2),
         col = guide_legend(ncol=2)) +
  theme_classic() +
  theme(legend.key.size = unit(5, 'mm'),
        legend.text = element_text(size = 6))

plot_pir
```

```{r combining_stages_data}
combined_data <- big_tpm %>% 
  rownames_to_column(var = 'Geneid') %>% 
  melt(value.name = 'tpm') %>% 
  mutate(stage = str_extract(variable, 
                             pattern = '(?<=_).+(?=_)')) %>% 
  group_by(Geneid, stage) %>% 
  dplyr::summarise(tpm_mean = mean(tpm),
                   tpm_med = median(tpm),
                   tpm_sd = sd(tpm),
                   tpm_max = max(tpm),
                   tpm_min = min(tpm))

combined_data <- dcast(combined_data, 
                       Geneid ~ stage, 
                       value.var = 'tpm_mean') %>% 
  column_to_rownames(var = 'Geneid')

#Ordering by life cycle progression
combined_data <- combined_data[,
                               ordering(colnames(combined_data), 
                                        order_vector = gsub(order_vector,
                                                            pattern = '-',
                                                            replacement = '\\.')
                               )]

#Check that they still sum to 1e6

apply(combined_data, 2, sum)

```

```{r colours_clades_phase}
SorL_col <- setNames(
  c(    '#F0E442',
        '#56B4E9',
        '#E69F00'),
  sort(
    unique(pir_clade %>% 
             mutate(
               SorL = str_extract(clade, 
                                  pattern = '(L|S(?=[[:digit:]]))|ancestral|pan-rodent')
             ) %>% 
             .$SorL)
  )
)

SorL_col_num <- SorL_col
names(SorL_col_num) <- sapply(names(SorL_col), 
                              function(clade_name) {
                                paste0(
                                  clade_name, 
                                  ' (n = ',
                                  sum(pir_clade$SorL == clade_name),
                                  ')'
                                )
                              })

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9',
    '#009E73',
    '#0072B2',
    '#E69F00',
    '#D55E00',
    '#CC79A7',
    '#999999',
    'black',
    'red'),
  sort(
   unique(pir_clade$clade)
  )
)

clade_col_num <- clade_col
names(clade_col_num) <- sapply(names(clade_col), 
                               function(clade_name) {
                                 paste0(
                                   clade_name, 
                                   ' (n = ',
                                   sum(pir_clade$clade == clade_name),
                                   ')'
                                 )
                               })
```

# pir expression, S or L-form, stacked bar chart, stack for each gene

```{r SorL_stacked_eachGene_fig2a}
  tpm_data <- combined_data %>% 
    rownames_to_column(var = 'Geneid') %>% 
    filter(Geneid %in% pir_id) %>% 
    melt(id.vars = 'Geneid', 
         variable.name = 'stage', 
         value.name = 'tpm') %>% 
  mutate(clade = pir_clade$clade[match(Geneid, 
                                       pir_clade$Geneid)]
         ) %>% 
  group_by(clade) %>% 
  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
    mutate(stack_group = Geneid) %>% 
    mutate(tpm = ifelse(tpm > 1, tpm, 0)) %>% #Filtering step for low value pir genes.
    group_by(stage, 
             stack_group,
             clade) %>% 
    dplyr::summarise(total_tpm = sum(tpm)) %>%
  arrange(
          desc(clade),
          total_tpm
)

tpm_data <- mutate(tpm_data,
                   order = row_number())

tpm_data_SorL_stackgenes <- mutate(tpm_data,
                        SorL = str_extract(clade, 
                                           pattern = '(L|S(?=[[:digit:]]))|ancestral|pan-rodent'))
tpm_data_SorL_stackgenes <- group_by(tpm_data_SorL_stackgenes,
                                     stage, 
                                     SorL,
                                     stack_group) %>% 
  dplyr::summarise(total_tpm = sum(total_tpm)) %>% 
  #Include the number of genes in each fill category for the labels
  mutate(SorL = paste0(SorL, 
                       ' (n = ', 
                       n_distinct(stack_group),
                       ')')) %>% 
#Order by SorL and then by total_tpm, then convert stack_group to a factor so ggplot orders correctly
  arrange(SorL, 
          total_tpm) %>% 
  mutate(stage_num = as.integer(stage))
tpm_data_SorL_stackgenes$stage %>% as.integer %>% unique
tpm_data_SorL_stackgenes$stage %>% unique

tpm_data_SorL_stackgenes$stack_group_order <- seq_len(length.out = nrow(tpm_data_SorL_stackgenes))
```

```{r stacked_bar_chart_figure2b}
#Calculate the range for the y value to make it look nicer
stacked_max <- tpm_data %>% 
  ungroup %>% 
  group_by(stage) %>% 
  dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
  .$max_tpm %>% 
  max

max_tpm <- c(
  seq(
    0,
    mround(
      stacked_max,
      500),
    length.out = 11)
)

stack_label <- 'sub-family'

write_csv(x = tpm_data_SorL_stackgenes,
          path = 'tl20-05_SorL-subfamily_pir.csv')

stage_vec <- unique(
  levels(tpm_data$stage)
)

stage_vec <- tpm_data_SorL_stackgenes$stage
col_group <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'zzAsex' ,
                   str_detect(stage_vec, pattern = 'Liv') ~ 'zzLiv',
                   str_detect(stage_vec, pattern = 'Gam') ~ 'zzGam',
                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'zzMos')
tpm_data_SorL_stackgenes$stage_col <- col_group
fills <- c('#F6B9AD','#96d3ff','#f0d4e3','#a7e9af') #Note that this order is influenced by sort(col_group)

y <- ggplot(data = tpm_data_SorL_stackgenes, 
            aes(
              x = stage_num,
              y = total_tpm, 
              fill = SorL,
              # group = stage_num
            )) +
  geom_rect(
    aes(
      xmin = stage_num - .7,
      xmax = stage_num + .7,
      ymin = -Inf,
      ymax = Inf,
      fill = stage_col
    ),
    alpha = .05
  ) +
  geom_bar(position = position_stack(),
           stat = 'identity',
           colour = 'black',
           size = 0.1) + 
  scale_fill_manual(
    breaks = unique(tpm_data_SorL_stackgenes$SorL),
    values = c(
      SorL_col_num,
      fills
    )
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  # scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) +
  scale_x_continuous(
    breaks = tpm_data_SorL_stackgenes$stage_num %>% unique(),
    labels = tpm_data_SorL_stackgenes$stage %>% unique(),
    expand = c(0, 0)
  ) +
    theme(
      axis.text.x = element_text(angle = 90, 
                                 vjust = .5,
                                 size = 10),
      axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    ) +
    xlab('stage') + 
    facet_grid(.~stage, 
               scales = "free_x",
               space = "free_x") +
    scale_x_discrete(expand=c(0,0)) +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x = element_blank(),
      # axis.line.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 10),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=8, 
                                  face = 'bold', 
                                  angle = 90),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      ),
      panel.spacing.x = unit(0, "null")
    )

g <- ggplot_gtable(ggplot_build(y))
strip_t <- which(grepl('strip-t', g$layout$name))
stage_vec <- unique(
  levels(tpm_data$stage)
)
fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
                   str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
                   str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
k <- 1
for (i in strip_t) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}

grid.newpage()
grid.draw(g)

```

# figure2binset

```{r stack_bar_graph_SorL_lowexpression-figure2binset}
#Include all stages expressed lower than a certain amount

low_express <- tpm_data_SorL_stackgenes %>% 
  group_by(stage) %>% 
  summarise(sum_tpm = sum(total_tpm)) %>% 
  filter(sum_tpm < 350)

#Calculate the range for the y value to make it look nicer

stacked_max <- max(low_express$sum_tpm)

max_tpm <- c(
  seq(
    0,
    mround(
      stacked_max,
      100),
    length.out = 11)
)

tpm_data_SorL_stackgenes_low <- filter(tpm_data_SorL_stackgenes,
                                       stage %in% low_express$stage)

y <- ggplot(data = tpm_data_SorL_stackgenes_low, 
            aes(
              x = stage,
              y = total_tpm, 
              fill = SorL,
              # group = stage_num
            )) +
  geom_bar(position = position_stack(),
           stat = 'identity',
           colour = 'black',
           size = 0.1) + 
  scale_fill_manual(
    breaks = unique(tpm_data_SorL_stackgenes$SorL),
    values = c(
      SorL_col_num,
      fills
    )
  ) +
  scale_color_manual('black') + 
  labs(fill = stack_label) +
  scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~stage, 
             scales = "free_x",
             space = "free_x") +
  scale_x_discrete(expand=expand_scale(add=1)) +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x = element_blank(),
    # axis.line.x = element_blank(),
    axis.text.y = element_text(size = 15),
    axis.title.y = element_text(size = 8),
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 5),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    strip.text.x = element_text(size=6, 
                                face = 'bold', 
                                angle = 90),
    strip.background = element_rect(
      color="black", 
      size=0.1, 
      linetype="solid"
    ),
    panel.spacing.x = unit(0, "null")
  )

#This bit is to change the colour of the facet boxes

g <- ggplot_gtable(ggplot_build(y))
strip_t <- which(grepl('strip-t', g$layout$name))
stage_vec <- unique(
  levels(tpm_data_SorL_stackgenes_low$stage)[levels(tpm_data_SorL_stackgenes_low$stage) %in% as.character(tpm_data_SorL_stackgenes_low$stage)]
)
fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
                   str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
                   str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
k <- 1
for (i in strip_t) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid.newpage()
grid.draw(g)
```

```{r table_of_SorL_highest_supptable}
#For a neat table
SorL_kable_index <- rep.int(2, 
                            times = length(unique(tpm_data_SorL_stackgenes$stage)))
names(SorL_kable_index) <- levels(tpm_data_SorL_stackgenes$stage)

tpm_data_SorL_stackgenes_top_table <- tpm_data_SorL_stackgenes %>% 
  mutate(Geneid = stack_group) %>% 
  filter(str_detect(SorL, 
                    pattern = 'ancestral', 
                    negate = TRUE)) %>% 
  group_by(stage) %>% 
  mutate(sum_tpm = sum(total_tpm)) %>% 
  group_by(stage, 
           SorL) %>% 
  slice_max(order_by = total_tpm,
            n = 1) %>% 
  mutate(total_tpm = round(total_tpm, digits = 2),
         prop_tpm = round(total_tpm/sum_tpm, digits = 2)) %>% 
  select(stage, Geneid, SorL, total_tpm, prop_tpm)

tpm_data_SorL_stackgenes_top_table %>% 
  kable() %>% 
  kable_styling() %>% 
  pack_rows(index = SorL_kable_index) %>% 
  scroll_box(width = "100%", height = "500px")
```

## Figure 2B - sub-family proportions

```{r combined_tpm_stacked_df}

  tpm_data_subfam <- combined_data %>% 
    rownames_to_column(var = 'Geneid') %>% 
    filter(Geneid %in% pir_id) %>% 
    melt(id.vars = 'Geneid', 
         variable.name = 'stage', 
         value.name = 'tpm') %>% 
  mutate(clade = pir_clade$clade[match(Geneid, 
                                       pir_clade$Geneid)]
         ) %>% 
  group_by(clade) %>% 
  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
    mutate(stack_group = clade) %>% 
    group_by(stage, 
             stack_group,
             clade) %>% 
    dplyr::summarise(total_tpm = sum(tpm)) %>%
  arrange(
          desc(clade),
          total_tpm
)

tpm_data_subfam <- mutate(tpm_data_subfam,
                   order = row_number())

  # #Apply ordering
  # tpm_data_subfam$stage = factor(tpm_data_subfam$stage, 
  #                         levels = ordering(tpm_data_subfam$stage,
  #                                           order_vector))

  #Calculate the range for the y value to make it look nicer
  stacked_max <- tpm_data_subfam %>% 
    ungroup %>% 
    group_by(stage) %>% 
    dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
    .$max_tpm %>% 
    max

  max_tpm <- c(
    seq(
      0,
      mround(
        stacked_max,
        500),
      length.out = 11)
  )
```

```{r combined_tpm_stacked_plot}

  stack_label <- 'sub-family'

y <- ggplot(data = tpm_data_subfam, 
            aes(
              x = stage,
              y = total_tpm, 
              fill = clade,
              group = order
            )) +
  geom_bar(
    # position = position_stack(),
    position = 'fill',
    stat = 'identity',
    colour = 'black',
    size = 0.1) +
  scale_fill_manual(
    values = clade_col_num
  ) +
  scale_color_manual('black') + 
  labs(fill = stack_label) +
  scale_y_continuous(breaks = seq(0,1,0.1)) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('Proportion of '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~stage, 
             scales = "free_x",
             space = "free_x") +
  scale_x_discrete(expand=expand_scale(add=1)) +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x = element_blank(),
    # axis.line.x = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    strip.text.x = element_text(size=8, 
                                face = 'bold', 
                                angle = 90),
    strip.background = element_rect(
      color="black", 
      size=0.1, 
      linetype="solid"
    ),
    panel.spacing.x = unit(0, "null")
  )

  #This bit is to change the colour of the facet boxes

  g <- ggplot_gtable(ggplot_build(y))
  strip_t <- which(grepl('strip-t', g$layout$name))
  stage_vec <- unique(

    levels(tpm_data_subfam$stage)
  )
  fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
                     str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
                     str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
                     str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
  k <- 1
  for (i in strip_t) {
    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
    k <- k+1
  }

  grid.newpage()
  grid.draw(g)
```

## Gametocyte marker gene expression - supp figure 2

```{r}
list_genes <- c(
  list('P28' = 'PBANKA_0514900'),
  list('nek4' = 'PBANKA_0616700'),
  list('map2' = 'PBANKA_0933700'),
  list('hap2' = 'PBANKA_1212600')
)
gam_markers <- unlist(list_genes)

list_tpms_line <- lapply(expt_tpms, 
                         function(input){
                           tpm_data <- input$tpm %>% 
                             filter(Geneid %in% gam_markers) %>% 
                             mutate(Geneid = names(gam_markers)[match(Geneid, gam_markers)]) %>% 
                             melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
                             mutate(stage = unlist(lapply(strsplit(as.character(sample), 
                                                                   split = '_'),
                                                          `[`,
                                                          1))) %>% 
                             group_by(stage, 
                                      sample,
                                      Geneid) %>% 
                             dplyr::summarise(total_tpm = sum(tpm))

                           #Make stages with no cirs in the sample give a zero 
                           if(!all(colnames(input$tpm) %in% tpm_data$sample)){
                             missing_stages = colnames(input$tpm)[-1][!colnames(input$tpm)[-1] %in% tpm_data$sample]
                             missing_stages_data_zero = data.frame(lapply(missing_stages, function(missing_st){
                               data.frame(stage = unlist(lapply(strsplit(as.character(missing_st), split = '_'),`[`,2)),
                                          sample = missing_st,
                                          total_tpm = 0)
                             }))
                             tpm_data <- bind_rows(tpm_data, missing_stages_data_zero)
                           }

                           #Add experiment column
                           tpm_data$experiment = rep(input$experiment, times = nrow(tpm_data))
                           max_tpm = if(max(tpm_data$total_tpm) < 10){
                             c(seq(0,mround(max(tpm_data$total_tpm),1),1))
                           }else{
                             c(seq(0,mround(max(tpm_data$total_tpm),5),5))
                           }

                           #Get the average data for the line graph
                           tpm_data_line <- tpm_data %>% 
                             ungroup %>% 
                             group_by(stage, 
                                      experiment,
                                      Geneid) %>% 
                             dplyr::summarise(total_tpm_median = median(total_tpm),
                                              Number_of_replicates = n(),
                                              total_tpm_max = max(total_tpm),
                                              total_tpm_min = min(total_tpm)) %>% 
                             arrange(stage)

                           return(tpm_data_line)
                         })

tpm_data_line <- bind_rows(list_tpms_line)

tpm_data_line$stage <- factor(tpm_data_line$stage, 
                             levels = ordering(tpm_data_line$stage,
                                               order_vector))

#Calculate the range for the y value to make it look nicer
if(max(
  tpm_data_line$total_tpm_max
) >= 500){
  max_tpm = c(
    seq(
      0,
      mround(
        max(
          tpm_data_line$total_tpm_max
        ),
        500),
      length.out = 11)
  )
}else{
  max_tpm = c(
    seq(
      0,
      mround(
        max(
          tpm_data_line$total_tpm_max
        ),
        10),
      length.out = 11)
  )
}
```

super hacky way to change the scales in each facet from:
https://fishandwhistle.net/post/2018/modifying-facet-scales-in-ggplot2/

```{r}
scale_override <- function(which, scale) {
  if(!is.numeric(which) || (length(which) != 1) || (which %% 1 != 0)) {
    stop("which must be an integer of length 1")
  }

  if(is.null(scale$aesthetics) || !any(c("x", "y") %in% scale$aesthetics)) {
    stop("scale must be an x or y position scale")
  }

  structure(list(which = which, scale = scale), class = "scale_override")
}

CustomFacetWrap <- ggproto(
  "CustomFacetWrap", FacetWrap,
  init_scales = function(self, layout, x_scale = NULL, y_scale = NULL, params) {
    # make the initial x, y scales list
    scales <- ggproto_parent(FacetWrap, self)$init_scales(layout, x_scale, y_scale, params)

    if(is.null(params$scale_overrides)) return(scales)

    max_scale_x <- length(scales$x)
    max_scale_y <- length(scales$y)

    # ... do some modification of the scales$x and scales$y here based on params$scale_overrides
    for(scale_override in params$scale_overrides) {
      which <- scale_override$which
      scale <- scale_override$scale

      if("x" %in% scale$aesthetics) {
        if(!is.null(scales$x)) {
          if(which < 0 || which > max_scale_x) stop("Invalid index of x scale: ", which)
          scales$x[[which]] <- scale$clone()
        }
      } else if("y" %in% scale$aesthetics) {
        if(!is.null(scales$y)) {
          if(which < 0 || which > max_scale_y) stop("Invalid index of y scale: ", which)
          scales$y[[which]] <- scale$clone()
        }
      } else {
        stop("Invalid scale")
      }
    }

    # return scales
    scales
  }
)

facet_wrap_custom <- function(..., scale_overrides = NULL) {
  # take advantage of the sanitizing that happens in facet_wrap
  facet_super <- facet_wrap(...)

  # sanitize scale overrides
  if(inherits(scale_overrides, "scale_override")) {
    scale_overrides <- list(scale_overrides)
  } else if(!is.list(scale_overrides) || 
            !all(vapply(scale_overrides, inherits, "scale_override", FUN.VALUE = logical(1)))) {
    stop("scale_overrides must be a scale_override object or a list of scale_override objects")
  }

  facet_super$params$scale_overrides <- scale_overrides

  ggproto(NULL, CustomFacetWrap,
    shrink = facet_super$shrink,
    params = facet_super$params
  )
}

(
  y <- ggplot(data = tpm_data_line, 
              aes(x = stage, 
                  y = total_tpm_median, 
                  fill = experiment, 
                  group = experiment)) +
    geom_bar(position = position_dodge2(preserve = 'single'),
             stat = 'identity') +
    scale_fill_manual(values = c(cbPalette, 'red', 'purple')[1:(length(unique(tpm_data_line$experiment)))]) +
    geom_errorbar(aes(x = stage,
                      ymax = total_tpm_max, 
                      ymin = total_tpm_min),
                  position = position_dodge2(preserve = 'single'),
                  size = 0.25) +
    scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_text(angle = 90, 
                                   size = 6, 
                                   vjust = 0.5),
          axis.text.y = element_text(size = 5),
          axis.title.y = element_text(size = 8),
          legend.title = element_text(size = 7),
          legend.text = element_text(size = 5),
          legend.key.size = unit(3,'mm'),
          legend.key.width = unit(2,'mm')) +
    scale_size(name = 'Number of genes detected over threshold', 
               range = c(0,3)) +
    ylab(
      bquote('total '~italic(.('pir'))~' TPM')
    ) +
    facet_wrap_custom(~Geneid, 
                      scales = 'free_y',
                      scale_overrides = list(
                        scale_override(3, scale_y_continuous(breaks = seq(0,650,length.out = 11))),
                        scale_override(2, scale_y_continuous(breaks = seq(0,900,length.out = 11))),
                        scale_override(1, scale_y_continuous(breaks = seq(0,310,length.out = 11)))
                      )) +
    annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)
)
#Could also use functions from a package called 'lemon' to add the lines to the top two facets.
```

## Lower expression of _pir_ genes versus the rest of the genome

```{r low_pir_expression}
set.seed(4908)

gene_expression_df <- combined_data %>% 
    rownames_to_column(var = 'Geneid') %>% 
    melt(id.vars = 'Geneid', 
         variable.name = 'stage', 
         value.name = 'tpm') %>% 
  mutate(pir_gene = ifelse(Geneid %in% pir_id,
                           'pir',
                           'non-pir')) %>% 
  group_by(stage, pir_gene) %>% 
  slice_sample(n = length(pir_id)) %>%
  summarise(mean_tpm = mean(tpm),
            med_tpm = round(median(tpm),
                            digits = 6))

g <- ggplot(gene_expression_df,
       aes(x = pir_gene,
           y = log2(1+med_tpm))) +
  xlab('Gene group')+
  ylab('l0g(Median TPM + 1)')+
  # geom_violin()+
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               # binwidth = 1,
               # dotsize = 2,
               method = 'dotdensity') +
  geom_path(aes(group = stage,
                col = stage)) +
  theme_classic() +
  theme(
    legend.text = element_text(size = 9),
    legend.key.size = unit(4, 'mm')
  )

g
```

_pir_ expression on average ranges from minimum mean `r min(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$mean_tpm)` TPM and median `r min(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$med_tpm)` TPM to maximum mean `r max(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$mean_tpm)` TPM and median `r max(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$med_tpm)` TPM. Non-\ _pir_ expression ranges from minimum mean `r min(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$mean_tpm)` TPM and median `r min(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$med_tpm)` TPM to maximum mean `r max(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$mean_tpm)` TPM and median `r max(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$med_tpm)` TPM.